Automated Verification and Synthesis of Embedded Systems: Advances and Challenges

### 2017-02-27 paper.tex
additions until section II

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, and consumer electronics. Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory and sophisticated software modules, in order to meet increasing computational power demands. As a consequence, reliability of embedded (distributed) software becomes a key issue during system development, which must be carefully addressed and assured. Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation. Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, and supporting new software architectures and legacy designs (usually written in low-level languages). The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems. Reliability issues, in the development of micro-grids and cyber-physical systems, are then considered, as a prominent verification and synthesis application for achieving a \textit{correct-by-construction} design.
\end{abstract}

@@

Generally, embedded computer systems perform dedicated functions with a high degree of reliability. They are used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}. Embedded systems are ubiquitous in modern day information systems and are also becoming increasingly important in our society, especially in micro-grids, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}, in multipurpose mobile devices, which now help in performing many routine daily tasks, and in cyber-physical systems (CPS), which demand short development cycles and again high levels of reliability~\cite{leeCPS,leeCPS2}. As a consequence, human life has also become more and more dependent on services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality/reliability.

Figure~\ref{intelligent-product} shows some examples of embedded systems, which typically consist of a human-machine interface ({\it e.g.}, keyboard and LCD), a processing unit ({\it e.g.}, real-time computer system), and an instrumentation interface ({\it e.g.}, sensor, network, and actuator) \cite{Kopetz11}. Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks. For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may cost civilian lives, which is unacceptable. In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health, with serious consequences.
%
\begin{figure}[!t]
	\centering
	\includegraphics[scale=0.35]{figure1.eps}
	\caption{An embedded system is part of a well-specified larger system (intelligent product).}
	\label{intelligent-product}
\end{figure}

Besides, when physical interaction with the real world is needed, which happens in CPS, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving. Regarding the latter, even human-in-the-loop feedback control can be employed \cite{munir}, which then raises deeper concerns related to the reliability of human behavior modeling and system implementation. Consequently, it is important to go beyond design correctness and also address behavior correctness, which may be performed by incorporating system models. Specifically, these models can be used to synthesize a given particular system, ensuring that all needed functions are correctly implemented and the correct behavior exhibited, {i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.

A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy. Additionally, the increasing computational power and decreasing size and cost, which are common to the area of computer processors, are enabling system designers to move more features to software domain, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification~\cite{Kroening15}.

We cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, and counter-example guided inductive synthesis, in Section~\ref{Preliminaries}. In Section~\ref{Verification-Challenges}, we describe the main challenges regarding the verification and synthesis of embedded systems, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research and Section \ref{Newapp} tackles new applications. Section~\ref{achievements} describes current achievements and future trends in verification and synthesis of embedded systems, while we conclude and describe future work in Section~\ref{conclusions}.

@@

In general, BMC techniques aim to check the violation of a given (safety) property at a given system depth, as shown in Figure~\ref{bounded-model-checking}. Indeed, given a transition system \textit{M}, which is derived from the control-flow graph of a program, a property $\phi$, which represents the program correctness and/or the system's behavior, and an iteration bound \textit{k}, which limits the loop unrolling, data structures, and context-switches, BMC techniques thus unfold the system \textit{k} times, in order to convert it into a verification condition $\psi$, expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.

@@

Nonetheless, there is still little evidence, in the available literature, that model checking hardware and software systems, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded-system verification. That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes. Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants about memory and variable aliasing~\cite{Bradley07}.

@@

The correctness specification $\sigma$ provided to the program synthesizer is of the form $\exists \vec{F} .  \forall \vec{x}.  \sigma(\vec{x}, \vec{F})$ where $\vec{F}$ ranges over functions, $\vec{x}$ ranges over ground terms, and $\sigma$ is a quantifier-free formula typically supported by SMT solvers. The ground terms is interpreted over some finite domain $\mathcal{D}$, where $\mathcal{D}$ can encoded using the SMT solvers bit-vectors theory. The {\sc Synthesize} and {\sc Verify} phases interact via a finite set of test vectors {\sc inputs} that is updated incrementally. Given the correctness specification $\sigma$, the {\sc Synthesize} procedure tries to find an existential witness $\vec{F}$ satisfying the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}$ in {\sc inputs} (as opposed to all $\vec{x} \in \mathcal{D}$). If {\sc synthesize} succeeds in finding a witness~$\vec{F}$, this witness is a candidate solution to the full synthesis formula, which is passed to the {\sc verify} phase to check whether it is a full solution ({\it i.e.}, $\vec{F}$ satisfies the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}\in\mathcal{D}$). If this is the case, then the algorithm terminates; otherwise, additional information is provided to the {\sc synthesize} phase in the form of a new counterexample that is added to the {\sc inputs} set and the loop iterates again. 
%
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.12]{figure4.eps}
	\caption{Counter-Example Guided Inductive Synthesis.}
	\label{Counter-Example-Guided-Inductive-Synthesis}
\end{figure}

@@

Nowadays, program synthesis engines that implement CEGIS approach~\cite{sketch} can automatically produce solutions for a large variety of specifications due to the combination of automated testing, genetic algorithms, and SMT-based automated reasoning~\cite{Sharma14}.

@@

Indeed, research in software verification is now incorporating such considerations, during checking processes, and some schemes already use knowledge about the system to be verified and the underlying hardware. Recently, a verification tool for digital systems was proposed, which is called digital system verifier (DSVerifier)~\cite{dsv_spin2015,Monteiro16} and is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length (FWL) effects. Additionally, DSVerifier checks closed-loop systems with uncertain models considering FWL effects, which are typically represented as hybrid systems, {\it i.e.}, the controller is digital but the controlled agent (plant) is a physical and continuous system~\cite{Bessa17}. Here, the verification procedure has to consider the interaction between a continuous plant and a digital (and sampled) controller with FWL effects, which can be connected using different control system configurations. Indeed, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions to explore the design space. Ultimately, if the system requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.

@@

@@

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The dependency on the correct functioning of embedded systems is rapidly growing, mainly due to their wide range of applications, such as micro-grids, automotive device control, health care, surveillance, mobile devices, and consumer electronics. Their structures are becoming more and more complex and now require multi-core processors with scalable shared memory and sophisticated software modules, in order to meet increasing computational power demands. As a consequence, reliability of embedded (distributed) software becomes a key issue during system development, which must be carefully addressed and assured. Normally, state-of-the-art verification methodologies for embedded systems generate test vectors (with constraints) and use assertion-based verification and high-level processor models, during simulation. Nonetheless, other additional challenges have been raised: the need for meeting time and energy constraints, handling concurrent software, dealing with platform restrictions, evaluating implementation-structure choices, validating operation logic, ensuring correct behavior and compliance with the target system, and supporting new software architectures and legacy designs (usually written in low-level languages). The present research discusses challenges, problems, and recent advances to ensure correctness and timeliness regarding embedded systems. Reliability issues, in the development of micro-grids and cyber-physical systems, are then considered, as a prominent verification and synthesis application for achieving a \textit{correct-by-construction} design.
\end{abstract}

@@

Generally, embedded computer systems perform dedicated functions with a high degree of reliability, according to their original design and requirements ({\it e.g.}, real time). They are ubiquitous in modern day information systems and are also becoming increasingly important in our society, due to their use to process, monitor, and control every human activity: factories, power plants, mobile devices, robotics, traffic control, vehicles, and home appliances. Besides, they are also used in a variety of sophisticated applications, which range from entertainment software, such as games and graphics animation, to safety-critical systems, such as nuclear reactors and automotive controllers~\cite{Kopetz11}.

The interaction between embedded software and physical processes created a different class of systems, which are complex, highly integrated, and present a mixture of continuous and discrete dynamics, {\it i.e.}, \textit{hybrid systems}. Indeed, the embedded-system domain allied to the most recent communication revolution caused an intense integration between different and spread physical processes (that are usually embedded systems), which are called cyber-physical systems (CPSs).

The main difference between embedded systems and CPSs is indeed the strong presence of modern communication technologies ({\it e.g.}, Internet of Things) in the last one. This simple difference causes a revolution in terms of flexibility, scalability, and complexity of those systems and adds a novel class of problems to the already known challenges of embedded systems. For example, micro-grids, {\it i.e.}, small-scale electricity systems that gather different sources of distributed generation (DG) and loads, are emerging CPSs, where reliability and carbon emission reduction are of paramount importance~\cite{xu15}. One may also notice that each DG element and the majority of current loads already employ embedded software and present several safety-critical requirements; however, the integration of both constitute a CPS, where additional challenges related to synchronization, stability, control, communication, and reliability of the entire micro-grid arise.

Thus, CPSs demand short development cycles and high level of reliability and robustness~\cite{leeCPS,leeCPS2}, besides presenting challenges that include but are not limited to the ones already imposed by embedded systems. As a consequence of their popularization, human life has also become more and more dependent on services provided by this type of system and, in particular, their success is strictly related to both service relevance and quality/reliability.

Figure~\ref{intelligent-product} shows some examples of embedded systems, which typically consist of a human-machine interface ({\it e.g.}, keyboard and LCD), a processing unit ({\it e.g.}, real-time computer system), and an instrumentation interface ({\it e.g.}, sensor, network, and actuator) \cite{Kopetz11}. Indeed, many current embedded systems, such as unmanned aerial vehicles (UAVs)~\cite{groza2015formal} and medical monitoring systems~\cite{Cordeiro09}, become interesting solutions only if they can reliably perform their target tasks. For instance, UAVs are a trend on military missions, due to the absence of pilots; however, an incorrect plan execution may cost civilian lives, which is unacceptable. In addition, wrong disease diagnosis or condition-evaluation reports have the potential to compromise patients' health and even compromise many aspects of their lives, with serious consequences. One the one hand, portable embedded systems are capable of monitoring and identifying conditions that are very difficult to a physician, mainly when their contact with patients happen only in medical clinics. On the other hand, wrong or diagnostic data may delay necessary treatments or compromise patients' health. 
%
\begin{figure}[!t]
	\centering
	\includegraphics[scale=0.35]{figure1.eps}
	\caption{An embedded system is part of a well-specified larger system (intelligent product).}
	\label{intelligent-product}
\end{figure}

Besides, when physical interaction with the real world is needed, which happens in CPSs, additional care must be taken, mainly when human action is directly replaced, as in vehicle driving. Regarding the latter, even human-in-the-loop feedback control can be employed \cite{munir}, which then raises deeper concerns related to the reliability of human behavior modeling and system implementation. Consequently, it is important to go beyond design correctness and also address behavior correctness, which may be performed by incorporating system models. Specifically, such models can be used to synthesize a given particular system, ensuring that all needed functions are correctly implemented and the correct behavior exhibited, {i.e.}, the system is indeed correct by its method of construction~\cite{Abate17}.

For instance, one may argue that a Linux device driver \cite{ldd} was verified and correctly uses the related kernel services; however, not much can be said about correct access to the associated hardware and its use, {\it i.e.}, if read and write operations are performed according to what is specified in the device's datasheet and setup periods are respected. In summary, the main goal would be ensuring that the developed driver is compliant with the underlying hardware, which goes further than code correctness.

A number of distinctive characteristics might influence the embedded system development and verification process, which include: mass production and static structure, functionality determined by software in read-only memory, multi-core processors with scalable shared memory, and limited amount of energy. Additionally, the increasing computational power and decreasing size and cost, which are common to the area of computer processors, are enabling system designers to move more features to software domain, which consequently leads to difficulties in verifying design correctness, since stringent constraints imposed by the underlying hardware ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency) must be considered during verification~\cite{Kroening15}. Such observations expand the addressed issues and even complement what was mentioned in the last paragraph.

We cover preliminaries on bounded model checking, \textit{k}-induction verification based on invariant inference, and counter-example guided inductive synthesis, in Section~\ref{Preliminaries}. In Section~\ref{Verification-Challenges}, we describe the main challenges regarding the verification and synthesis of embedded systems, while Section~\ref{Research-Problem} tackles those challenges that are (partially) open in current published research. Section \ref{Newapp} then tackles new applications, which take into account the new paradigm addressed here, and Section~\ref{achievements} describes current achievements and future trends in verification and synthesis of embedded systems. Finally, we conclude and describe future work in Section~\ref{conclusions}.

@@

In general, BMC techniques aim to check the violation of a given (safety) property at a given system depth, as shown in Figure~\ref{bounded-model-checking}. Indeed, given a transition system \textit{M}, which is derived from the control-flow graph of a program, a property $\phi$, which represents the program correctness and/or the system's behavior, and an iteration bound \textit{k}, which limits the loop unrolling, data structures, and context-switches, BMC techniques thus unfold a system \textit{k} times, in order to convert it into a verification condition $\psi$, expressed in propositional logic or in a decidable-fragment of first-order logic, such that $\psi$ is \textit{satisfiable} if and only if $\phi$ has a counterexample of depth less than or equal to \textit{k}.

@@

Nonetheless, there is still little evidence, in the available literature, that model checking hardware and software systems through \textit{k}-induction (and invariants) can be efficiently exploited in embedded-system verification. That happens due to the distinctive characteristics mentioned earlier, which influence the embedded system development and also the employed verification processes. Additionally, there is still a lack of studies for embedded software verifiers to exploit the combination of different invariant generation and strengthening algorithms, including analysis to discover linear inequalities, polynomial equalities and inequalities, and invariants about memory and variable aliasing~\cite{Bradley07}.

@@

The correctness specification $\sigma$ provided to the program synthesizer is of the form $\exists \vec{F} .  \forall \vec{x}.  \sigma(\vec{x}, \vec{F})$ where $\vec{F}$ ranges over functions, $\vec{x}$ ranges over ground terms, and $\sigma$ is a quantifier-free formula typically supported by SMT solvers. The ground terms is interpreted over some finite domain $\mathcal{D}$, where $\mathcal{D}$ can be encoded using the SMT solvers bit-vectors theory. The {\sc Synthesize} and {\sc Verify} phases interact via a finite set of test vectors {\sc inputs} that is updated incrementally. Given the correctness specification $\sigma$, the {\sc Synthesize} procedure tries to find an existential witness $\vec{F}$ satisfying the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}$ in {\sc inputs} (as opposed to all $\vec{x} \in \mathcal{D}$). If {\sc synthesize} succeeds in finding a witness~$\vec{F}$, this witness is a candidate solution to the full synthesis formula, which is passed to the {\sc verify} phase to check whether it is a full solution ({\it i.e.}, $\vec{F}$ satisfies the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}\in\mathcal{D}$). If this is the case, then the algorithm terminates; otherwise, additional information is provided to the {\sc synthesize} phase in the form of a new counterexample that is added to the {\sc inputs} set and the loop iterates again. 
%
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.12]{figure4.eps}
	\caption{Counter-Example Guided Inductive Synthesis.}
	\label{Counter-Example-Guided-Inductive-Synthesis}
\end{figure}

@@

Nowadays, program synthesis engines that implement CEGIS approach~\cite{sketch} can automatically produce solutions for a large variety of specifications, due to the combination of automated testing, genetic algorithms, and SMT-based automated reasoning~\cite{Sharma14}.

@@

Indeed, research in software verification is now incorporating such considerations, during checking processes, and some schemes already use knowledge about the system to be verified and the underlying hardware. Recently, a verification tool for digital systems was proposed, which is called digital system verifier (DSVerifier)~\cite{dsv_spin2015,Monteiro16} and is able to aid engineers to check overflow, limit cycle, output error, timing, stability, and minimum phase, considering finite word-length (FWL) effects. Additionally, DSVerifier checks closed-loop systems with uncertain models considering FWL effects, which are typically represented as hybrid systems, {\it i.e.}, the controller is digital but the controlled agent (plant) is a physical and continuous system~\cite{Bessa17}. That ultimately leads to the use of analog-to-digital converters, which are one of the most important aspects to be considered, given that data loss (quantization) is inevitable. As a consequence, the verification procedure has to consider the interaction between a continuous plant and a digital (and sampled) controller with FWL effects, which can be connected using different control system configurations. Additionally, the latter may present a different behavior, regarding what was specified in analog domain, due to the inherent discret-time operation. Indeed, DSVerifier is a useful test tool, which takes into account different representations ({\it e.g.}, transfer-function and state-space), realization forms ({\it e.g.}, direct forms, delta forms, and transposed forms) and other implementation restrictions, in order to explore the design space. finally, if a system's requirements are not met with a given configuration, an analysis of the provided error report may suggest another setup.

@@

\section{New Applications: Beyond Code Correctness}
\label{Newapp}

@@

\bibitem{ldd}
Witkowski T., Blanc N., Kroening D., Weissenbacher G.:
\newblock {\em {Model checking concurrent linux device drivers}}.
\newblock In: 22nd IEEE/ACM international conference on Automated software engineering, pp. 501--504, 2007.
